---
title: "Compiliing the final dataset"
format: 
    html:
      css: cayman.css

editor: source
---

# Setting up the working environment

## Required packages

```{r}
#| warning: false
#| message: false

library(raster)
library(tidyverse)
library(rnaturalearth)
library(cowplot)
library(sf)
library(plyr)
library(Utilities.Package)
library(classInt)
library(fpc)
library(BBmisc)
library(performance)
library(mgcv)
```

## Folder structures

Various folders where the data is stored (if working on work Windows laptop):

```{r}
# wd<-"/Volumes/One Touch/Main_folder/In_prep_submittted_manuscripts/Bridges_etal_OBIS_assessment/OBIS_data_clean/Code/"

wd<-"/Volumes/One Touch/Main_folder/In_prep_submittted_manuscripts/Bridges_etal_OBIS_assessment/OBIS_data_clean/Code/"

final_benthic_datasets_folder<-paste0(wd, "step9_final_datasets/Benthic/")

final_pelagic_datasets_folder<-paste0(wd, "step9_final_datasets/Pelagic/")

# gebco_folder<-"/Volumes/One Touch/Main_folder/GIS Files/World_Layers/GEBCO/GEBCO_2020/"

gebco_folder<-"/Volumes/One Touch/Main_folder/GIS Files/World_Layers/GEBCO/GEBCO_2020/"

```

## Generate list of desired columns

```{r}
# cols_keep<-c("decimalLatitude", "decimalLongitude", "gebco_depth_value","depth","obis_record_depth","predicted_b_min","marine","brackish", "dataset_id","kingdom","phylum","subphylum", "class","subclass","order","suborder", "family","genus","species", "scientificName", "aphiaID", "BP")
```

## Read in data

```{r}
# section1<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section1_270623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section2a<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section2a_270623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section2b<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section2b_280623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section3<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section3_280623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section4<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section4_300623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section5<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section5_300623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section6<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section6_300623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section7<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section7_300623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
# section8<-read.csv(paste0(final_benthic_datasets_folder, "Individual_sections/GEBCO_vals_TaxonID2_section8_300623_pelagic.csv"), head=T, sep=",")%>%
#   dplyr::select(all_of(cols_keep))
```

## Combine & remove non-marine taxa

```{r}
# full_df<-bind_rows(section1, section2a, section2b, section3, section4, section5, section6, section7, section8)
# 
# marinefalse<-full_df %>%
#   filter(marine=="FALSE")
```

**Benthic**
Some of these supposedly false records are probably correct and should be marine=TRUE. I'm going to eyeball the dataset at the class level instead. 

**Pelagic**


```{r}
# class_list<-as.data.frame(unique(full_df$class))
```

**Benthic**
Of the 84 results, NA is one, two are non-marine groups (Diplopoda and Aves) - I have removed these.

**Pelagic**
Of the 84 results, NA is one, one is non-marine group (Aves) and one doesn't come up on WoRMS (Inarticulata) - I have removed these.

```{r}
# full_df<-full_df %>%
#   filter((class!="Aves") %>%
#   replace_na(TRUE))
# 
# full_df<-full_df %>%
#   filter((class!="Inarticulata") %>%
#   replace_na(TRUE))

```

Write out our final df. 

```{r}
# write.csv(full_df, paste0(final_datasets_folder, "AllBenthicRecords_TaxonID2_Global_FINAL.csv"), row.names = F)
# write.csv(full_df, paste0(final_datasets_folder, "AllPelagicRecords_TaxonID2_Global_FINAL.csv"), row.names = F)

full_df<-read.csv(paste0(final_benthic_datasets_folder, "AllBenthicRecords_TaxonID2_Global_FINAL.csv"), head=T, sep=",")
full_df_p<-read.csv(paste0(final_pelagic_datasets_folder, "AllPelagicRecords_TaxonID2_Global_FINAL.csv"), head=T, sep=",")

# full_df_new<-rbind(full_df_b, full_df)
```

## Map

```{r}
# world_coordinates <- map_data("world") 
# 
# map<-ggplot() + 
#   geom_map(data = world_coordinates, map = world_coordinates, aes(long, lat, map_id = region))+
#   geom_point(data=full_df, 
#              aes(x=decimalLongitude, y=decimalLatitude), colour="darkturquoise", 
#              fill="darkgrey",pch=21, size=0.5, alpha=I(0.7))

# full_df<-full_df%>%
#   mutate(hemisphere=case_when(decimalLatitude < 0 ~ "S",
#                               decimalLatitude > 0 ~ "N"))
# 
# check<-full_df %>%
#   group_by(hemisphere) %>%
#   summarise(sum=count(hemisphere))

```

## Get important summary info

### Group by phylum

```{r}
# phylum_records<-full_df %>%
#   drop_na(phylum)
# 
# phylum_records_p<-full_df_p %>%
#   drop_na(phylum)

```


### Split vertebrates and invertebrates

Firstly, remove the 32,215 records with no phylum.

```{r}
# check<-phylum_records %>%
#   group_by(phylum) %>%
#   summarise(sum=count(phylum))
# 
# check_p<-phylum_records_p %>%
#   group_by(phylum) %>%
#   summarise(sum=count(phylum))
#
# write.csv(check_p, paste0(final_pelagic_datasets_folder, "pelagic_phylum_summary.csv"), row.names = F)
```

```{r}

# inverts<-phylum_records %>%
#   filter((subphylum!="Vertebrata") %>%
#            replace_na(TRUE))%>%
#   filter((subphylum!="Cephalochordata") %>%
#            replace_na(TRUE))
# 
# verts<-phylum_records %>%
#   filter(subphylum=="Vertebrata" |
#          subphylum=="Cephalochordata")
  
```

We do need to consider records where subphylum is NA but the phylum is Chordata. For these records, we don't know if they're inverts or verts. 

```{r}

# non_chordate_inverts<-inverts %>% 
#   filter(phylum!="Chordata")
# 
# chordata_inverts<-inverts %>%
#   filter(phylum=="Chordata")
# 
# check<-chordata_inverts %>% 
#   group_by(subphylum) %>% 
#   summarise(sum=count(subphylum))

```

There are 2,000 records like this. Let's remove them. 

```{r}
# keep<-chordata_inverts %>% 
#   drop_na(subphylum)
# 
# final_inverts<-rbind(non_chordate_inverst, keep)
# 
# write.csv(final_inverts, paste0(final_datasets_folder, "AllBenthicRecords_Invertebrates_Global_FINAL.csv"), row.names = F)

```

Make a shapefile. 

```{r}
# final_inverts<-read.csv(paste0(final_datasets_folder, "AllBenthicRecords_Invertebrates_Global_FINAL.csv"), sep=",", head=T)
# 
# final_inverts_shp<-st_as_sf(final_inverts, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)%>%
#   dplyr::rename(b_min=predicted_b_min,
#          gebco_dep=gebco_depth_value,
#          obis_dep=obis_record_depth,
#          sciName=scientificName)
# 
# 
# output_shapefile<-paste0(final_datasets_folder, "AllBenthicRecords_Invertebrates_Global_FINAL.shp")
# 
# st_write(final_inverts_shp, output_shapefile)

```

### No. invertebrate records to at least genus

```{r}
# genus_records<-final_inverts %>% 
#   group_by(genus) %>% 
#   summarise(sum=count(genus))
```

4,022,075 invertebrate genus records.

## Filter regionally if needed

```{r}
north_atlantic<-full_df %>% 
  filter(decimalLatitude >= 5 &
           decimalLatitude <= 69 &
           decimalLongitude <= 15 &
           decimalLongitude >= -100)

write.csv(north_atlantic, "D:/OneDrive/OneDrive - University of Plymouth/In_prep_submittted_manuscripts/Allcock_etal_C150NAReview/GapAnalysisData/CheckedVersions/RoughNA_Pelagic_OBIS.csv", row.names = F)
```

## Convert df to shp

```{r}
full_shp<-st_as_sf(full_df, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)%>%
  dplyr::rename(b_min=predicted_b_min,
         gebco_dep=gebco_depth_value,
         obis_dep=obis_record_depth,
         sciName=scientificName)


output_shapefile<-paste0(final_datasets_folder, "AllBenthicRecords_TaxonID2_Global_FINAL.shp")

st_write(full_shp, output_shapefile)

```

## Convert data to raster grid

```{r}
# gebco_tile<-raster(paste0(gebco_folder,"GEBCO2020_depth_430m_WGS84.tif")) 
# gebco_tile<-raster(paste0(gebco_folder,"GEBCO2020_depth_1km_WGS84.tif"))
# gebco_tile<-raster(paste0(gebco_folder,"GEBCO2020_depth_10km_WGS84.tif"))
gebco_tile<-raster(paste0(gebco_folder,"GEBCO_1deg_Template_WGS84.tif"))

coordinates(full_df) <- c("decimalLongitude", "decimalLatitude")

point_count_raster <- rasterize(full_df, gebco_tile, field = 1, fun = "count")

writeRaster(point_count_raster, paste0(final_benthic_datasets_folder,"AllBenthicRecords_TaxonID2_Global_1deg_FINAL.tif"), format = "GTiff")

```

